Probabilistic Forecasts in Hierarchical~Time~Series

### 2018-07-06 ProbabilisticCoherence.tex
Cleanup of probabilistic reconciliation

We now extend the methodology of point forecast reconciliation to probabilistic forecasts

%For probabilistic forecasts, reconciliation implies finding the probability measure of the coherent forecasts using the information from an incoherent probabilistic forecast measure. A more formal definition is given below.

  Let $(\mathbb{R}^n, \mathscr{F}_{\mathbb{R}^n}, \hat{\nu})$ be an probability triple, that is incoherent and that characterises forecast uncertainty for all variables in the hierarchy at time $t+h$ conditional on all information up to time $t$.  This may be obtained from the first stage of the forecasting process e.g. by modelling and forecasting each series individually.  Let $g:\mathbb{R}^n \rightarrow \mathbb{R}^m $ be a linear mapping.  Let $(\mathbb{R}^m, \mathscr{F}_{\mathbb{R}^m}, \nu)$ be a probability triple defined on $\mathbb{R}^m$.
  \begin{definition} \label{def:reconprob}
  We define the reconciled probability measure of $\hat{\nu}$ with respect to the mapping $g(.)$ as a probability measure $\tilde{\nu}$ on $\mathfrak{s}$ with $\sigma$-algebra $\mathscr{F}_\mathfrak{s}$ where the following holds
  \begin{equation}
  \tilde{\nu}(s(\mathcal{B})) = \nu(\mathcal{B})= \hat{\nu}(g^{-1}(\mathcal{B})) \qquad \forall \mathcal{B} \in \mathscr{F}_{\mathbb{R}^m}\,,
  \end{equation}
  where $g^{-1}(\mathcal{B}):=\left\{\breve{\bm{y}}\in \mathbb{R}^n:g(\breve{\bm{y}})\in \mathcal{B} \right\}$ is the pre-image of $\mathcal{B}$, that is the set of all points in $\mathbb{R}^n$ that $g(.)$ maps to a point in $\mathcal{B}$.
\end{definition}

@@

\begin{figure}
	\input{Figs/probforerec_schematic.tex}
	\caption{Summary of probabilistic forecast reconciliation.  The probability that ${\bm y}_{t+h|t}$ lies in the red line segment under the reconciled probabilistic forecast is defined to be equal to the probability that ${\bm y}_{t+h|t}$ lies in the shaded blue area under the unreconciled probabilistic forecast.  Note that since the smallest hierarchy involves three dimensions, this figure is only a schematic}\label{fig:probfr_sch} 
\end{figure}

@@

While $\hat{\nu}$ is a probability measure for an $n$-vector $\hat{\bm{y}}_{t+h|t}$,  probability statements in terms of a different coordinate system can be made via an appropriate change of basis. Letting $f(.)$ be generic notation for a probability density functions and following the notation from our definition of point forecast reconciliation where $\hat{\bm{y}}_{t+h|t}=\bm{S}\tilde{\bm{b}}_{t+h|t}+\bm{R}\tilde{\bm{a}}_{t+h|t}$ we obtain

@@

After a change in basis

\begin{equation}
f(\tilde{\bm b},\tilde{\bm a})=(2\pi)^{-\frac{n}{2}}\Big|\bm{\hat{\Sigma}}_{t+h}\Big|^{-\frac{1}{2}}\Big|(\bm{S} ~  \bm{R})\Big|\exp\left\{-\frac{1}{2}q\right\}\,,
\end{equation}
where
\begin{equation}
q=(\bm{S}\tilde{\bm{b}}+\bm{R}\tilde{\bm{a}}-\bm{\hat{\mu}})' \bm{\hat{\Sigma}}^{-1}(\bm{S}\tilde{\bm{b}}+\bm{R}\tilde{\bm{a}}-\bm{\hat{\mu}})
\end{equation}

@@

%Equation \eqref{ex:2.1} implies $\tilde{\bm{b}}_{t+h} \sim \mathcal{N}(\bm{P}\bm{\hat{\mu}}_{t+h}, \bm{P}\hat{\bm{\Sigma}}_{t+h}\bm{P}')$, where $\bm{P} = (\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$. Then from \eqref{4.7} it follows that
%\begin{equation}\label{eq:gaussianreconciled}
%\tilde{\bm{f}}(\tilde{\bm{y}}_{t+h})=\tilde{\bm{f}}(\bm{S}\tilde{\bm{b}}_{t+h}).
%\end{equation}
Which implies that the reconciled probabilistic forecast for the bottom level series is $\tilde{\bm{b}}_{t+h} \sim \mathcal{N}(\bm{G}\bm{\hat{\mu}}_{t+h}, \bm{G}\hat{\bm{\Sigma}}_{t+h}\bm{G}')$.  The reconciled probabilistic forecasts for the whole hierarchy follow a degenerate Gaussian distribution with mean  $\bm{SG}\bm{\hat{\mu}}$ and rank deficient covariance matrix $\bm{SG}\hat{\bm{\Sigma}}_{t+h}\bm{G}'\bm{S}'$.
\subsection{Elliptical Distributions}
We now show that the true predictive distribution can be recovered for elliptical distributions via linear reconciliation.  Here, for any square matrix ${\bm C}$, ${\bm C}^{1/2}$ and ${\bm C}^{-1/2}$ are defined to satisfy ${\bm C}^{1/2}\left({\bm C}^{1/2}\right)'={\bm C}$ and ${\bm C}^{-1/2}\left({\bm C}^{-1/2}\right)'={\bm C}^{-1}$, for example ${\bm C}^{1/2}$ may be obtained via the Cholesky or eigenvalue decompositions. 
\begin{theo}[Reconciliation for Elliptical Distributions]
	Let an unreconciled probabilistic forecast come from the elliptical class with location parameter $\hat{\bm \mu}$ and scale matrix $\hat{\bm \Sigma}$.  Let the true predictive distribution of ${\bm y}_{t+h|t}$ also belong to the elliptical class with location parameter ${\bm \mu}$ and scale matrix ${\bm \Sigma}$.  Then the affine reconciliation mapping $g(\breve{\bm y})={\bm G}_{opt}\breve{\bm y}+{\bm d}_{opt}$ with $\bm{G}_{opt}={\bm A}{\bm\Sigma}^{-1/2}$ and ${\bm d}_{opt}={\bm \mu}-{\bm S}{\bm G}_{opt}\hat{\bm \mu}$ recovers the true predictive density where ${\bm A}$ is any $m\times n$ matrix such that ${\bm A}{\bm A}'={\bm \Omega}$ and ${\bm \Omega}$ is the true variance covariance matrix of the predictive distribution for the bottom level. 
\end{theo}
\begin{proof}
   Since elliptical distributions are closed under affine transformations, and are closed under marginalisation, reconciliation of an elliptical distribution yields an elliptical distribution (although the unreconciled and unreconciled distributions may be different members of the class of elliptical distributions).  The scale matrix of the reconciled forecast is given by $\bm{S}\bm{G}_{opt}{\bm \Sigma}{\bm G}_{opt}'{\bm S}'$ while the location matrix is given by $\bm{S}\bm{G}_{opt}\hat{\bm \mu}+{\bm d}_{opt}$.  The reconciled scale matrix is 
   \begin{align*}
   \tilde{\bm \Sigma}_{opt}=&{\bm S}{\bm A}{\bm\Sigma}^{-1/2}{\bm \Sigma}\left({\bm\Sigma}^{-1/2}\right)'{\bm A}'{\bm S}'\\
   =&{\bm S}{\bm \Omega}{\bm S}'\\
   =&{\bm \Sigma}
   \end{align*}
   For the choices of ${\bf G}_{opt}$ and ${\bf d}_{opt}$ given above, the reconciled location vector is 
   \begin{align*}
   \tilde{\bm \mu}_{opt}=&{\bm S}{\bm G}_{opt}\hat{\bm \mu}+{\bm \mu}-{\bm S}{\bm G}_{opt}\hat{\bm \mu}\\
   =&{\bm \mu}\\
   \end{align*}
\end{proof}	
A number of insights can be drawn from this theorem.  First, although a linear mapping $g(.)$ can be used to recover the true density in the elliptical case, the same does not hold in general.  Second, $g(.)$ is not, in general, a projection matrix.  The conditions for which a the true predictive density can be recovered by reconciliation are given below.

\begin{theo}[True predictive via projection]
	Assume that the true predictive distribution is elliptical with location ${\bm mu}$ and scale ${\bm Sigma}$.  Consider reconciliation via a projection $g({\bm y})=({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}{\bm y}$
	The true predictive distribution can be recovered via reconciliation of an elliptical distribution with location $\hat{\bm \mu}$ $\hat{\bm \Sigma}$ and scale when the following conditions hold.
	\begin{align}
	sp(\hat{\bm\mu}-{\bm \mu})&\subset sp(\bm{R})\\
    sp(\bm{\Sigma}^{1/2}-{\bm \Sigma}^{1/2})&\subset sp({\bm R})\\
	\end{align}
\end{theo}
\begin{proof}
	The reconciled location vector will be given by
	\begin{align*}
		\tilde{\bm \mu}&={\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}\hat{\bm \mu}\\
		&={\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}\left(\hat{\bm \mu}+{\bm \mu}-{\bm \mu}\right)\\
		&={\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}{\bm \mu}+{\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}\left(\hat{\bm \mu}-{\bm \mu}\right)\\
	\end{align*}
  Since ${\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}$ is a projection onto $\mathfrak{s}$ and ${\bm \mu}\in\mathfrak{s}$ the first term simplified to ${\bm \mu}$.  If ${\bm \mu}-\hat{\bm \mu}$ lies in the span of ${\bm R}$ then multiplication by ${\bm R}'_{\perp}$ reduced the second term to ${\bm 0}$.  By a similar argument it can be shown that $\tilde{\bm \Sigma}^{1/2}={\bm \Sigma}^{1/2}$.  The closure property of elliptical distributions under affine transformations ensures that the full true predictive distribution can be recovered.
\end{proof}  
Although these conditions will rarely hold in practice and only apply to a limited class of distributions they do provide some insight into selecting a projection for reconciliation.  If the value of $\hat{\bm \mu}$ were equi-probable in all directions then a projection orthogonal to $\mathfrak{s}$ would be a sensible choice for ${\bm R}$ since it would in some sense represent a `median' direction for ${\bm \mu}-\hat{\bm \mu}$.  However, the one step ahead in-sample errors are usually correlated suggesting that $\hat{\bm \mu}$ is more likely to fall in some directions than others.  As such an orthogonal projection after transformation by the inverse of the one step ahead in-sample errors may be more intuitively appealing.  This is exactly what the MinT projection estimates, and as simulations will show in Section~\ref{sec:gaussian}, this projection leads to the best empirical results.

>>>>

We now extend the methodology of point forecast reconciliation to probabilistic forecasts
%For probabilistic forecasts, reconciliation implies finding the probability measure of the coherent forecasts using the information from an incoherent probabilistic forecast measure. A more formal definition is given below.
Let $(\mathbb{R}^n, \mathscr{F}_{\mathbb{R}^n}, \hat{\nu})$ be an probability triple, that is incoherent and that characterises forecast uncertainty for all variables in the hierarchy at time $t+h$ conditional on all information up to time $t$.  This is obtained from the first stage of the forecasting process e.g. by modelling and forecasting each series individually.  Let $g:\mathbb{R}^n \rightarrow \mathbb{R}^m $ be a linear mapping.  Let $(\mathbb{R}^m, \mathscr{F}_{\mathbb{R}^m}, \nu)$ be a probability triple defined on $\mathbb{R}^m$.
  \begin{definition} \label{def:reconprob}
  The reconciled probability measure of $\hat{\nu}$ with respect to the mapping $g(.)$ is a probability measure $\tilde{\nu}$ on $\mathfrak{s}$ with $\sigma$-algebra $\mathscr{F}_\mathfrak{s}$ where the following holds
  \begin{equation}
  \tilde{\nu}(s(\mathcal{B})) = \nu(\mathcal{B})= \hat{\nu}(g^{-1}(\mathcal{B})) \qquad \forall \mathcal{B} \in \mathscr{F}_{\mathbb{R}^m}\,,
  \end{equation}
  where $g^{-1}(\mathcal{B}):=\left\{\breve{\bm{y}}\in \mathbb{R}^n:g(\breve{\bm{y}})\in \mathcal{B} \right\}$ is the pre-image of $\mathcal{B}$, that is the set of all points in $\mathbb{R}^n$ that $g(.)$ maps to a point in $\mathcal{B}$.
\end{definition}

@@

\begin{figure}
	\input{Figs/probforerec_schematic.tex}
	\caption{Summary of probabilistic forecast reconciliation.  The probability that ${\bm y}_{t+h}$ lies in the red line segment under the reconciled probabilistic forecast is defined to be equal to the probability that ${\bm y}_{t+h}$ lies in the shaded blue area under the unreconciled probabilistic forecast.  Note that since the smallest hierarchy involves three dimensions, this figure is only a schematic}\label{fig:probfr_sch} 
\end{figure}

@@

While $\hat{\nu}$ is a probability measure for an $n$-vector $\hat{\bm{y}}_{t+h|t}$,  probability statements in terms of a different coordinate system can be made via an appropriate change of basis. Letting $f(.)$ be generic notation for a probability density function and following the notation from our definition of point forecast reconciliation where $\hat{\bm{y}}_{t+h|t}=\bm{S}\tilde{\bm{b}}_{t+h|t}+\bm{R}\tilde{\bm{a}}_{t+h|t}$ we obtain

@@

In an alternative basis

\begin{equation}
f(\tilde{\bm b},\tilde{\bm a})=(2\pi)^{-\frac{n}{2}}\Big|\bm{\hat{\Sigma}}\Big|^{-\frac{1}{2}}\Big|(\bm{S} ~  \bm{R})\Big|\exp\left\{-\frac{1}{2}q\right\}\,,
\end{equation}
where
\begin{equation}
q=(\bm{S}\tilde{\bm{b}}+\bm{R}\tilde{\bm{a}}-\bm{\hat{\mu}})' \bm{\hat{\Sigma}}^{-1}(\bm{S}\tilde{\bm{b}}+\bm{R}\tilde{\bm{a}}-\bm{\hat{\mu}})
\end{equation}

@@

%Equation \eqref{ex:2.1} implies $\tilde{\bm{b}}_{t+h} \sim \mathcal{N}(\bm{P}\bm{\hat{\mu}}_{t+h}, \bm{P}\hat{\bm{\Sigma}}_{t+h}\bm{P}')$, where $\bm{P} = (\bm{R}'_\bot \bm{S})^{-1}\bm{R}'_\bot$. Then from \eqref{4.7} it follows that
%\begin{equation}\label{eq:gaussianreconciled}
%\tilde{\bm{f}}(\tilde{\bm{y}}_{t+h})=\tilde{\bm{f}}(\bm{S}\tilde{\bm{b}}_{t+h}).
%\end{equation}
Which implies that the reconciled probabilistic forecast for the bottom level series is $\tilde{\bm{b}}_{t+h} \sim \mathcal{N}(\bm{G}\bm{\hat{\mu}}_{t+h}, \bm{G}\hat{\bm{\Sigma}}_{t+h}\bm{G}')$.  The reconciled probabilistic forecasts for the whole hierarchy follow a degenerate Gaussian distribution with mean  $\bm{SG}\bm{\hat{\mu}}$ and rank deficient covariance matrix $\bm{SG}\hat{\bm{\Sigma}}_{t+h}\bm{G}'\bm{S}'$.
\subsection{Elliptical Distributions}
We now show that the true predictive distribution can be recovered for elliptical distributions via linear reconciliation.  Here, for any square matrix ${\bm C}$, ${\bm C}^{1/2}$ and ${\bm C}^{-1/2}$ are defined to satisfy ${\bm C}^{1/2}\left({\bm C}^{1/2}\right)'={\bm C}$ and ${\bm C}^{-1/2}\left({\bm C}^{-1/2}\right)'={\bm C}^{-1}$, for example ${\bm C}^{1/2}$ may be obtained via the Cholesky or eigenvalue decompositions. 
\begin{theo}[Reconciliation for Elliptical Distributions]
	Let an unreconciled probabilistic forecast come from the elliptical class with location parameter $\hat{\bm \mu}$ and scale matrix $\hat{\bm \Sigma}$.  Let the true predictive distribution of ${\bm y}_{t+h|t}$ also belong to the elliptical class with location parameter ${\bm \mu}$ and scale matrix ${\bm \Sigma}$.  Then the linear reconciliation mapping $g(\breve{\bm y})={\bm G}_{opt}\breve{\bm y}+{\bm d}_{opt}$ with $\bm{G}_{opt}={\bm A}{\bm\Sigma}^{-1/2}$ and ${\bm d}_{opt}={\bm \mu}-{\bm S}{\bm G}_{opt}\hat{\bm \mu}$ recovers the true predictive density where ${\bm A}$ is any $m\times n$ matrix such that ${\bm A}{\bm A}'={\bm \Omega}$ and ${\bm \Omega}$ is a sub-matrix of ${\bm \Sigma}$ corresponding to the bottom level series. 
\end{theo}
\begin{proof}
   Since elliptical distributions are closed under affine transformations, and are closed under marginalisation, reconciliation of an elliptical distribution yields an elliptical distribution (although the unreconciled and reconciled distributions may be different members of the class of elliptical distributions).  The scale matrix of the reconciled forecast is given by $\bm{S}\bm{G}_{opt}{\bm \Sigma}{\bm G}_{opt}'{\bm S}'$ while the location matrix is given by $\bm{S}\bm{G}_{opt}\hat{\bm \mu}+{\bm d}_{opt}$.  The reconciled scale matrix is 
   \begin{align*}
   \tilde{\bm \Sigma}_{opt}=&{\bm S}{\bm A}{\bm\Sigma}^{-1/2}{\bm \Sigma}\left({\bm\Sigma}^{-1/2}\right)'{\bm A}'{\bm S}'\\
   =&{\bm S}{\bm \Omega}{\bm S}'\\
   =&{\bm \Sigma}
   \end{align*}
   For the choices of ${\bf G}_{opt}$ and ${\bf d}_{opt}$ given above, the reconciled location vector is 
   \begin{align*}
   \tilde{\bm \mu}_{opt}=&{\bm S}{\bm G}_{opt}\hat{\bm \mu}+{\bm \mu}-{\bm S}{\bm G}_{opt}\hat{\bm \mu}\\
   =&{\bm \mu}\\
   \end{align*}
\end{proof}	
A number of insights can be drawn from this theorem.  First, although a linear mapping $g(.)$ can be used to recover the true predictive in the elliptical case, the same does not hold in general.  Second, $g(.)$ is not, in general, a projection matrix.  The conditions for which a the true predictive density can be recovered by a projection are given below.

\begin{theo}[True predictive via projection]
	Assume that the true predictive distribution is elliptical with location ${\bm \mu}$ and scale ${\bm \Sigma}$.  Consider reconciliation via a projection $g({\bm y})=({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}{\bm y}$. The true predictive distribution can be recovered via reconciliation of an elliptical distribution with location $\hat{\bm \mu}$ and scale $\hat{\bm \Sigma}$  when the following conditions hold.
	\begin{align}
	sp(\hat{\bm\mu}-{\bm \mu})&\subset sp(\bm{R})\\
    sp(\hat{\bm \Sigma}^{1/2}-{\bm \Sigma}^{1/2})&\subset sp({\bm R})\\
	\end{align}
\end{theo}
\begin{proof}
	The reconciled location vector will be given by
	\begin{align*}
		\tilde{\bm \mu}&={\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}\hat{\bm \mu}\\
		&={\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}\left(\hat{\bm \mu}+{\bm \mu}-{\bm \mu}\right)\\
		&={\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}{\bm \mu}+{\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}\left(\hat{\bm \mu}-{\bm \mu}\right)\\
	\end{align*}
  Since ${\bm S}({\bm R}'_{\perp}{\bm S})^{-1}{\bm R}'_{\perp}$ is a projection onto $\mathfrak{s}$ and ${\bm \mu}\in\mathfrak{s}$ the first term simplifies to ${\bm \mu}$.  If ${\bm \mu}-\hat{\bm \mu}$ lies in the span of ${\bm R}$ then multiplication by ${\bm R}'_{\perp}$ reduces the second term to ${\bm 0}$.  By a similar argument it can be shown that $\tilde{\bm \Sigma}^{1/2}={\bm \Sigma}^{1/2}$.  The closure property of elliptical distributions under affine transformations ensures that the full true predictive distribution can be recovered.
\end{proof}  
Although these conditions will rarely hold in practice and only apply to a limited class of distributions they do provide some insight into selecting a projection for reconciliation.  If the value of $\hat{\bm \mu}$ were equi-probable in all directions then a projection orthogonal to $\mathfrak{s}$ would be a sensible choice for ${\bm R}$ since it would in some sense represent a `median' direction for ${\bm \mu}-\hat{\bm \mu}$.  However, the one step ahead in-sample errors are usually correlated suggesting that $\hat{\bm \mu}$ is more likely to fall in some directions than others.  As such an orthogonal projection after transformation by the inverse of the one step ahead in-sample errors may be more intuitively appealing.  This is exactly what the MinT projection provides, and as simulations will show in Section~\ref{sec:gaussian}, this projection leads to the best empirical results.
